---
title: "R Programming - Loops"
author: "Your Name"
date: '...'
output:
  word_document: default
  html_document: default
---


```{r setup, include=FALSE}
#Do not change anything in this chunk
options(width=150)
knitr::opts_chunk$set(echo = TRUE,collapse=TRUE)
library(regclass)
```

*********************

```{r load in saved enviroment}
#Make sure the .RData file is in your analytics folder along with this .Rmd file.  
#If you're unsure where the .Rmd file is, do File then Save As and manually save it there.
load("RProgrammingPractice.RData")  
```


****************

1.  (basic `for` loop) Although setting `x` to be equal to the sequence 5, 10, 17, ..., 101 is easy with `x <- (1 + 1:9)^2 + 1`, write a `for` loop that defines each element of `x` one at a time (the `i`-th element of `x` should be the value `(1+i)^2+1`.  Check that the first element of `x` is indeed 5 and your vector is of length 9 (with no `NA`s).  Be sure to initalize `x` before the loops!

```{r Q1}

```

Walkthrough:

a)  Since you're going to be placing elements into `x` one at a time, you need to make sure it's "initialized" first.

b)  Think about what you're looping over.  In this case it is "elements of `x`", or even more specifically "positions inside the vector `x`".  Call the looping variable `position`, let `position <- 3`, and get your code working so it defines the 3rd element of `x` correctly as the number `(1+3)^2+1` = 17 (but write your code using `position` instead of hard-coding 3!).

c)  Figure out how to write the vector of values to loop over (this vector will go to the right of the `in` inside the `for ( in ) {}`).

d)  Put it all together and generalize your line from (b) so that it'll work inside the `for` loop.  


****************

2.  (basic `for` loop) The infinite sequence 1 + 1/2 + 1/4 + 1/8 + 1/16 + ... + 1/`2^i` + ... can be shown to sum to exactly 2.  Verify that this seems reasonable by defining (using a `for` loop) a vector `x` whose elements are the first 50 values of this series.  Then, print to the screen the quantity 2 minus the sum of the elements of `x` (which should be really, really close to 0).  Note:  this is more easily accomplished by running `2 - sum( (1/2)^(0:49) )`, but it gives good practice!


```{r Q2}

```

Walkthrough:

a)  Since you're going to be placing elements into `x` one at a time, you need to make sure it's "initialized" first.

b)  Think about what you're looping over.  You are creating a vector whose 1st element is (1/2) raised to the 0 power, whose 2nd element is (1/2) raised to the 1st power, whose 3rd element is (1/2) raised to the 2nd power, etc.  The loop is over "elements of `x`" (since we are defining one at a time), or even more specifically "positions inside the vector `x`".  

Call the looping variable `i`, let `i <- 4`, and get your code working so it defines the 4rd element of `x` correctly as the `(1/2)^(4-1)` (1/2 raised to the third power), but write your code using `i` instead of hard-coding 4!).

c)  Figure out how to write the vector of values to loop over (this vector will go to the right of the `in` inside the `for ( in ) {}`).

d)  Put it all together and generalize your line from (b) so that it'll work inside the `for` loop.  



****************

3.  (`for` loop over categories)  Although this task is more easily accomplished by running `aggregate(TipPercentage~Weekday,data=TIPS,FUN=mean)` (find the average tip percentage for each day of the week), use a `for` loop to populate the elements of a named vector called `average.tip` with the average tip percentage for each day of the week in the `Weekday` column of the `TIPS` data from `regclass`.  You can verify the numbers by running the aforementioned `aggregate` command.

```{r Q3}
library(regclass)
data(TIPS)
aggregate(TipPercentage~Weekday,data=TIPS,FUN=mean)

#Try for Saturday

#Generalize

```

Walkthrough:  

a)  Run the aggregate command so you know what numbers to expect.

b)  Since you're going to be placing elements into `average.tip` one at a time, you need to make sure it's "initialized" first.

c)  Think about what you're looping over.  In this case it is "days of the week", or even more specifically "levels of `Weekday`", so `levels(TIPS$Weekday)` will be the vector of values to loop over.

d)  Let the looping variable be called `day`, and try out `day <- "Saturday"`. Develop code that finds the average tip percentage on Saturday (instead of hard-coding Saturday anywhere, let it be represented by `day`).  One way:  take a subset or rows that correspond to Saturday, then find the average of the `TipPercentage` column for that subset.

e)  Set up the `in` for the loop: `for ( in ) {}`.  You've named the looping variable and know the vector of values to loop over.

f)  Generalize your line from (d) so that it'll work inside the `for` loop.  Use the `which` trick to store the average in the appropriate position of the `average.tip` vector you have created.


***************

4.  (`for` loop over categories)  The dataframe `PRODUCT` in `regclass` has information on whether a new product was considered a success (`Outcome` column, levels are `success` and `fail`) and the product's category (`Category`, anonymized to be A, B, C, D).   Write a `for` loop that produces a vector called `probfailure` which contains the proportion of failures for each product category.  


```{r Q4}
data(PRODUCT)
#Let's see what numbers we are after
aggregate(Outcome~Category,data=PRODUCT,FUN=function(x)mean(x=="fail"))

#Vector of values that we'll be looping over
levels(PRODUCT$Category)

#Get it working for specific case

#Generalize

```

Tips:  

a)  Run  `aggregate(Outcome~Category,data=PRODUCT,FUN=function(x)mean(x=="fail"))`  so you know what numbers to expect (this command will make more sense when we talk about writing our own functions).

b)  Since you're going to be placing elements into `probfailure` one at a time, you need to make sure it's "initialized" first.

c)  Determine what you're looping over.  In this case it is "product categories", or even more specifically "levels of `Category`", so `levels(PRODUCT$Category)` will be the looping vector.

d)  Let the looping variable be called `product`, and try out `product <- "C"`. Develop code that finds the proportion of failures for product C (instead of hard-coding C anywhere, let the product be represented by `product`).  One way:  take a subset or rows that correspond to product C, then find the average value of the logical vector `Outcome=="fail"` for that subset (or count up the number of times "fail" is in the Outcome column and divide that by the number of rows in the subset).

e)  Set up the `in` for the loop: `for ( in ) {}`.  You've named the looping variable and know the vector of values to loop over.

f)  Generalize your line from (d) so that it'll work inside the `for` loop.  Use the `which` trick to store the proportion in the appropriate position of the `probfailure` vector you have created.


***************

5.  (`for` loop)  Write a for loop that goes through the length of the mtcars vector and adds together the mpg.  Start by creating a place to store values in to a vector.  Then add together the values to the previous value in a for loop.

```{r Q5}

```


****************


6.  Basic Lists. Run the model below and investigate the object ML. What is it's dimenstion? What is the third object in the list?  What is the value of the 3000th residual? Print the intercept to the screen using the ML list object. 

```{r  Q6}
## Run this code first
ML <- lm(V4~V6, data = NYSE)
summary(ML)

## Dimension?

## Thrid object? Use code, not just inspection. 

## 3000th residual

## Intercept


```

****************

7.  (basic `for` loop)   The command `sample(x,size=n,replace=TRUE)` picks n elements from the vector `x` at random with replacement (the same element can be picked twice).  Imagine that over a customer's lifetime they make either 1, 2, ..., or 10 purchases (for all intents and purposes the value can be considered to be picked at random).  Each purchase results in the customer spending either 5, 5.5, 6, 6.5, ..., or 20 dollars (for all intents and purposes the value can be considered to be picked at random).   

Using a `for` loop, generate the lifetime values (total spent) of 50000 customers and make a histogram of the results.  What is the overall average lifetime value?  You'll want to follow best practices for writing `for` loops to get this right!

Hints:

* Get your code working for the first customer before writing the loop!  

* First, create a variable called `purchases` which contains the (randomly picked) number of purchases that this customer makes during his or her lifetime (i.e., left-arrow into `purchases` the output of the appropriate `sample` command).

* Then, create a vector `amounts` by left-arrowing the output of the appropriate `sample` command.  This vector will store the amounts of money spent on each of this customers purchases.  For instance, if `purchases` happened to equal 6, then `amounts` will be a vector of length 6 containing 6 randomly chosen monetary values. If `purchases` happened to be 2, then this will be a vector of length 2 containing 2 randomly chosen monetary values.  The sum of the values in `amounts` gives this customer's lifetime value.

* If you need help setting up the `sample` command to generate `amounts`, go back to the description of `sample`.  The first argument is a vector of values to select from (use `seq` to specify this).  The second argument is the number of values to pick.  This number is "random", but is stored in `purchases`!  

* Once you are confident in your code, initialize a vector `lifetimevalue` to be an empty vector, and have a `for` loop fill in its values.  You're looping over the integers 1, 2, 3, (The i-th element of `lifetimevalue` will be the lifetime value of customer `i` and will equal the sum of the numbers in the `amounts` vector created).

```{r Q7}


```



************

8.  (basic `for` loop)  Consider the  `AcquisitionCost` vector in the global environment that is found after loading in the .RData file.  This vector records the effective amount of money a company had to spend to acquire each of 500 different customers.    Using a `for` loop, define the elements of a vector named `running.avg` such that that its first element is the first element of `AcquisitionCost`, its second element is the average of the first two elements of `AcquisitionCost`, its third element is the average of the first three elements of `AcquisitionCost`, etc.   Sanity check:  you should find that the 278th element of `running.avg` is 497.9221.  

*Print to the screen elements 353-361 of `running.avg` as well as the results of running `summary` on it*

```{r Q8}



```


**********

9.  (`for` loop over categories) Consider the `Income` vector in the global environment.  This is a 6-level categorical variable giving the approximate yearly incomes of 500 individuals.  Using a `for` loop, create a vector called `clvQ20` that stores the 20th percentile of the customer lifetime values (stored in the `LifetimeValue` vector in the global environment) for each level of income.  Then, make a dataframe whose first column (named `IncomeLevel`) contains the levels of `Income` and whose second column (called `Percentile20`) contains the corresponding percentiles.

Tip 1:  if `x` is a vector of numerical values, remember that `quantile(x,.20)` gives the 20th percentile.  If 155 happened to be the 20th percentile, that means 20% of lifetime values are smaller than 155 while 80% are larger.

Tip 2:  `Income` and `LifetimeValue` exist in the global environment and not in a data frame, so the syntax from most examples in class/lectures need to be slightly modified.  If you are more comfortable using the syntax had these been columns in a dataframe, then run `DATA <- data.frame(Income,LifetimeValue)` to create one.

*Print to the screen the contents of the requested dataframe*.  You can check your work by running `aggregate(LifetimeValue~Income,FUN=quantile,probs=.2)` which produces a similar dataframe (but with different column names).

```{r Q9}


```


*********

10.  Create an object based on the summary(ML) fron the example in 6 above. Query that object to make the t-value for the intercept print below this chunk. NOTE: you will use the chuck to test your ideas. Ultimately, the output is below the chunk. The chunk is suppressed. 

```{r Q10, include=FALSE}
## Create object here

## Find t-value here


## NOTE: the statement below should print with the t-value showing (you have to add something to make that happen.)
```
*********
**The t-value for the intercept is   .**

11b.  Customers all have unique shopping behaviors of course, but sometimes there are patterns  There might be a cluster of customers who spend most of their money on meat, another cluster who spends most of their money on the salad bar, etc.   Let's study the two departments where people spend most of their money.

* Let's start by examining the customer in row 143.  Left-arrow `DEPARTMENT[143,]` into a vector called `x`.  By looking at the contents of `x` (perhaps sorting it), it's clear that this customer spends the most money in `DELI` and second most in `GROCERY`. 

```{r Q11b first bullet}
x <- DEPARTMENT[143,]
sort(x)
```

* To figure how to automatically extract the names of the departments with the two highest value, print to the screen `sort(x,decreasing=TRUE)` and `names( sort(x,decreasing=TRUE) )`.  The category with the most money is the first element of the output of `names` and the category with the second most money is the second element of the output of `names`.  

```{r Q11b second bullet}
sort(x,decreasing=TRUE)
names( sort(x,decreasing=TRUE) )
names( sort(x,decreasing=TRUE) )[1]
names( sort(x,decreasing=TRUE) )[2]
```

*  Initialize vectors called `most` and `second` to be blank vectors.  We'll use a `for` loop to fill in its elements.  Specifically, the value in position `i` should be the department name ("FROZEN", "PRODUCE", "VIDEO RENTAL", etc.) that customer in row `i` has spent the most money (`most`) or second most money (`second`).  Based on what we've seen so far, we know that `most[143]` will be "DELI" and `second[143]` will be "GROCERY".  Verify that this is the case, and also print to the screen `most[222]` and `second[222]` (which should be "GROCERY" and "DRUG GM", respectively).

```{r Q11b third bullet}


```

* Once `most` and `second` have been defined, print to the screen the result of running `table(most)` and `table(second)` to get an idea of how often each department appears in these vectors.  Also, write a single line of code that prints to the screen the number of customers where both most equals "GROCERY" AND second equals "GAS". 

```{r Q11b fourth bullet}


```

****************
12. Plot the residuals versus the fitted values for the ML object. Put a horizontal line at zero. The graph will look weird. Remove the row that is an outlier and replot. 

```{r 12}

```

