---
title: "Simple Regression Wk 2"
author: "Grayson Pettit"
date: '2024-09-24'
output: word_document
---
Run this chunk first

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(320)
library(regclass)
```
Key pdf introduction pages: #  pdf Pg 7, 14, 17, 18, 21, 22, 27,


Play around with the otlier demo to see how a single point can impact a line. Put the code in the console, not in the .rmd.


outlier_demo()  #Pg34, 77

pg 37-38


In the chunk below, make 4 vectors, "x, y, z, w". x is a vector of length 50, where the first 25 entries are random (provided) and the next 25 are a carbon copy of the first 25. Let y = x^2 plus random normal variation. Let z = 50 random normal variates. Let w = z^2 plus random normal variation.

```{r repeat}
?length
x <- rnorm(n=25)
x= append(x,x, after = length(x))
x
y= x^2 +rnorm(n=50)
z= rnorm(n=50)
w= z^2 + rnorm(n=50)

# This gets you started. The final x must be twice as long

# final x

# y

# z

# w








```


Use the code below to explore the w,x,y and z vectors. Add your own code to check regression assumptions and visualize the model. Remove the #'s lines 52-55.


```{r explore, echo=FALSE}
plot(y~x)
M = lm(y~x)
plot(M)
summary(M)
abline(M,col="red")

visualize_model(M)
# Pg 46 checing regression


 # Pg 86 Review visualizing a model




```

Repeat the chunk above analyzing w ~ z

```{r}
# w ~ z
plot(w~z)
M1 = lm(w~z)
plot(M1)
summary(M1)
abline(M1,col="red")
visualize_model(M1)






```

Use the chunck below "check for transformations" in * model M *. You should find that the "best" model is y ~ x^2. Make a new variable xsq = x^2 and run all the plots and analyses on that model. 

```{r transform}

# Checking transformations:  pg 61




xsq = x^2
plot(y~xsq)
M2<-lm(y~xsq)
plot(M2)
summary(M2)
abline(M2,col="red")
visualize_model(M2)
find_transformations(M2, powers=seq(from=-3, to=3, by=.25), threshold = .3)
```

Practice predicting. Using the prior model y ~ xsq, predict the values of y for x = -5 , 0, 1, 3.  Get both confidence and prediction intervals. 

```{r predict}
TO.PREDICT<-data.frame(xsq= c(-5, 0, 1, 3))
predict(M2,newdata=TO.PREDICT,interval="confidence",level=0.90)



```

