---
title: "Diamond"
output: word_document
date: "2023-03-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Introduction to Multiple Regression (Part 1)

# Key pdf pages: 7-8, 15, 27, 42, 45*, 61-4, 66-7, 74-6, 81-4. 89, 108, 114-5

# New concept: Adjusted r-square. 

# New Functions:
    ##  VIF()
    ##  pairs()
    ##  anova(simple, complex)  sort of new

# New Argument:
    ## check_regression (M, extra = TRUE)
  
# Continued use:
    ## confint()
    ## predict()
    ## cor_matrix()

```{r}
library(regclass)
```




Read in Round Prices dataset. Set of diamond prices and characteristics. Call data.frame 'D'

MUST create D before trying to Knit ******************************************
```{r read in}
D = read.csv('~/Downloads/Major Course work: Materials/BAS 320 (Linear Regression) Grayson Pettit/7.1 Round Prices.csv')

M= lm(Price ~ Carat, data=D)

summary(M)

MR= lm(Price~., data = D)
summary(MR)

anova(MR)


# Elplore the datastet. 

```



Make a simple linear regression model 'M' with the most obvious x variable to predict Price (Y). If you don't know what the most obvious variable is, ask your instructor. Summarize the model. 

```{r Make first model}
M = lm(Price~Carat, data = D)

summary(M)


```
Interpret r-squre in the context of the problem. Not the official definition, but as you needing to buy a diamond.
We can explain 18.7% of variation within the slope of price between two different diamonds with the variable carat alone. 


Now make a model(MR) with all the variables, using "twidle dot". Sumarize the model.
```{r}
MR = lm(Price~., data = D)
check_regression(MR, extra = TRUE)
summary(MR)
anova(MR)
visualize_model(MR)
```

What in world is all that output??? What do you think of r-square?


Run this line of code (make sure your model name equals mine). What do you think the output is telling you?
```{r}
# anova(MR)
```




Run this code
```{r Ordinal}
Diamond = D[,c(1:4,6)]

Diamond$Cut = ifelse (D$Cut == "Good", 1, 
                ifelse(D$Cut == "Very Good", 2,
                       ifelse(D$Cut == "Excellent", 3,4))
)

hist(Diamond$Cut)

# Color
# D$Color = factor(D$Color)
# plot(D$Color)

Diamond$Color = ifelse (D$Color == "K", 1, 
                ifelse(D$Color == "J", 2,
                       ifelse(D$Color == "I", 3,
                              ifelse(D$Color == "H", 4,
                                     ifelse(D$Color == "G", 5,
                                             ifelse(D$Color == "F", 6,
                                                    ifelse(D$Color == "E",7,8))))))
)

# Now Clarity

 #D$Clarity =factor(D$Clarity)
# plot(D$Clarity) 

Diamond$Clarity = ifelse (D$Clarity == "SI1", 1, 
                  ifelse(D$Clarity == "SI2", 2,
                         ifelse(D$Clarity == "VS1", 3,
                                ifelse(D$Clarity == "VS2", 4,
                                       ifelse(D$Clarity == "VVS1", 5,
                                              ifelse(D$Clarity == "VVS2", 6,7
                                                     )))))
)





```


############# WHEWWWWWWWWWWWWWWWWWW That was a lot ###########################


##   now back to the story
Look at the correlations between all the variables in Diamond. 
```{r correlations}
###  pgs. 17,106



cor_matrix(Diamond)

```
 

Copy multiple regression code above, change D to Diamond. Run all the same analyses. 
```{r}
MR0= lm(Price~Carat, data = Diamond)

MR1 = lm(Price~., data = Diamond)
check_regression(MR1, extra = TRUE)
summary(MR1)
anova(MR1)
visualize_model(MR1)

VIF(MR1)

to.predict= data.frame(Carat = c(2,.5), Cut = c(3,1), Color = c(6,1), Clarity =c(5,2))
predict(MR1, newdata = to.predict, interval = "confidence", level= .95)

cor_matrix(Diamond)

pairs(Diamond)

pairs(Price~ Carat + Clarity, data = Diamond)

anova(MR0, MR1)

confint(MR1)
```


Run the appropriate test to see if adding variables other than Carat are statistically significant as a whole

```{r}
## hint M.simple Pg 113


```



Check the regression assumptions on the multiple regression model. Run confidence intervals on the estimates. 

```{r}

# hint extra = TRUE

```



You are loooking at two diamonds that you may buy. Estimate their price and CI for their average price. 
```{r predict}
# Diamond 1: Carat = 2 ,Cut = 3, Color = 6 ,Clarity = 5
# Diamond 2: Carat = .5 ,Cut = 1, Color = 1 ,Clarity = 2





```



