---
title: "Activity 3 - Associations Part 1"
author: "Grayson Pettit"
date: "..."
output:
  word_document: default
---


In this activity, and for the rest of the course, we will use `regclass` package.  
This package contains the commands written specifically for association and regression analysis and contains many datasets.

Once `regclass` is installed, load it up and make it available by running

```{r load up regclass, message=FALSE}
library(regclass)
```

**Question 1:**
After the `regclass` library has been loaded, load in the `SURVEY11` dataset from package `regclass` by running

```{r read SURVEY11}
data(SURVEY11)

```

This contains information about Stat 201 students from 2011.  Let us study a potential association between someone's relationship status `X34.RelationshipStatus` (the $y$ variable) and someone's gender `X02.Gender`.  If an association existed, this would imply that men and women view definitions of relationship statuses somewhat differently.  

Use `associate`, adding the argument `seed=320` so that the set of 500 (default value) permutation datasets generated are the same for everyone.

```{r}
#Your associate command using a seed of 320
associate(X34.RelationshipStatus ~ X02.Gender, data = SURVEY11, seed = 320)
```

- Does the mosaic plot suggest that an association exists?  Why or why not?

*Response:* 
- More men are single than women, this is statistically significant at the 5% level. 
**************

- What is the estimated $p$-value of the "discrepancy" between the conditional distributions of relationship status between genders?  Why is the test inconclusive?

*Response:*
- With 95% confidence, my p value is between .011 and .039, these numbers are less than p=.05, therefore there is an indication of association between gender and relationship status. However, this sample size is 628 individuals on a college campus, therefore we are not at the true beta due to small sample size. I would also assume that the standard errors are large and we cannot see that from the association function. 
**************

- Add the argument `permutations=1500` to make 1500 permutation datasets instead of the default 500 (have the seed still be 320).  The test will now be conclusive.  Is the association statistically significant?

```{r}
#Your associate command using 1500 permutations and a seed of 320

associate(X34.RelationshipStatus ~ X02.Gender, data = SURVEY11, permutations = 1500, seed = 320)

```

*Comment:*
- Yes, with 95% confidence my P-value is between .015 and .03

**************

**Question 2:**  
Download friendpotF.dat to your BAS320 folder (or whever this .Rmd document is stored) and read it into R using `read.csv`, calling it `FF`.  This is the dataset of friendship potentials (for women) generated by you guys, as well as attractiveness scores and other characteristics as evaluated by other classes.

```{r}
FF <- read.csv("friendpotF.dat")  
```

**************

- Look at a histogram and a QQ-plot of `FriendScore` using `hist()` and `qq()`.  Remember to refer to a specific column of the `FF` dataframe you will need to write it as `FF$FriendScore`.  Does the mean or does the median provide the better summary of the typical value of `FriendScore`?

```{r}
#R code for histogram and qqplot
```

*Comment:*

**************

- Determine if there is an association between friendship potential (`FriendScore`) and the sexuality of the person `Actual.Sexuality`.  Use a seed of 320.  Report the differences in averages (or medians if that was the appropriate measure) and the $p$-value of this difference along with a comment about statistical significance. 

```{r}
#associate code
```

*Comment:*

**************

- Go through some of the other categorical variables and determine if associations exist.  If an association exists, which level has the highest average friendship potential?   Possible choices:  `ApparentRace`, `Chin`, `Cleavage`, `ClothingStyle`, `Glasses`, `HairColor`, `LookingAtCamera`, `Selfie`, `Piercings`.  If a test is inconclusive with the default 500 permutations (e.g., `Glasses`), up the number of permutations.


```{r, warning=FALSE}
#associate(FriendScore~ApparentRace,data=FF,seed=320) #Uncomment and run
#many more lines using associate, one for each variable, all using seed 320
```

*Comment:*
