---
title: "BAS 471 - Unit 2 Parts 1 and 2 - Intro to Monte Carlo Simulations"
author: "Monte Marlo Mavens"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1 (Kahoot 1):  "for" loop in your head

Run the following "for" loop in your head. What is the value of `x` when the loop terminates?

```{r Q1}
# Given in class

x<- 0
values<-c(2,4,3)
for (i in 1:length(values)){
  x<- x + 1 + values[i]
  
}
x



```

## Question 2  (Kahoot 2):  How does sample work?

What will the following command output?

```{r Q2}
# Given in class
model<- "Petrie"
outfits<- c("hoodie", "buttons", "tshirts")

for(i in outfits){
  dress(model, i)
  strut(model)
  
}

```

**Response:**




## Question 3  (Kahoot 3):  How does sample work?

Why won't the following sample command work?

```{r Q3}
# Given in class


```

**Response:** 





## Question 4

Four integers between 1-20 are picked at random with replacement.  Consider two scenarios.  One where all numbers are all equally likely.  Another where the weight assigned to the integer `x` equals `5+x^2` (i.e. larger numbers are progressively more likely).  Via Monte Carlo simulations with 100,000 trials, estimate the following probabilities. 

a.  **Kahoot 4** What is the probability that the numbers picked will add to 63?  

```{r Q4a}
possible_values=1:20
weights<- 5+ possible_values^2
ntrials<- 100000
counter<-0

for( i in 1:ntrials){
value<-sample(possible_values, size=4, replace= TRUE, prob = weights ) 
  if( sum(value)==63){ counter<- counter+1}
}
counter/ntrials




```

b. **Kahoot 5** What is the probability that the largest number is greater than the sum of the smallest 2.  


```{r Q4b}

possible_values=1:20
weights<- 5+ possible_values^2
ntrials<- 100000
counter<-0

for( i in 1:ntrials){
value<-sample(possible_values, size=4, replace= TRUE, prob = weights ) 
  if( sum(sort(value)[1:2])<max(value)){ counter<- counter+1}
}
counter/ntrials

```



c.  What is the probability that all are even numbers.  I found about 8.2%.

```{r Q4c}


```

d.  At least 2 values are 15 or more.  I found about 85%.

```{r Q1d}


```


*************

## Question 5

Two contestants are playing a trivia game.  After a question is asked, each contestant has to "buzz in" before answering.  If the contestant wins the buzz in, and answers correctly, they win a point.  If they get it wrong, the game moves to the next question (the other contestant does not get a chance to answer).  Assume:

* Contestant A has a probability of 70% of answering correctly, but has a 40% chance of winning the buzz in.
* Contestant B has a probability of 50% of answering correctly, but has a 60% chance of winning the buzz in.

30 total questions are asked.  Using a Monte Carlo simulation with 10000 trials to estimate

* **Kahoot 6** the probability that contestant A wins (has a score strictly larger than B)?

* the probability that contestant B wins (has a score strictly larger than A)?

* the probability that game ends in a tie?

One approach:  for each question, select who wins the buzzer appropriately, then determine if that contestant got it right.  I defined variables `score.A` and  `score.B` to keep track of the scores each game, and incremented counters I had defined called `win.A`, `win.B`, `tie` appropriately after 30 questions had been asked.

Another approach:  Let `correct.A` be a vector of 30 1s and 0s to represent whether contestant A would answer correctly (had they won the buzzer).  Define `correct.B` similarly.  Then define a vector `winbuzz` to be a vector of 30 1s and 0s to represent who wins the buzz-in (1 represents A winning and 0 represents B winning).  What does `sum(winbuzz*correct.A)` tell us?  What `sum( (1-winbuzz)*correct.B)` tell us?

```{r Q5}



```


***********


### Question 6 (Kahoot 7)

Imagine a pitch has a 60% chance of being a strike and a 40% chance of being a ball.  A batter has a "strike out" if 3 strikes occur before 4 balls.  A batter has a "walk" if 4 balls occur before 3 strikes.

With a Monte Carlo simulation of 10,000 trials, estimate the probability that a batter will have a "strike out" (3 strikes occur before 4 balls). 

Hint:  let 1 represent a strike and 0 represent a ball.  Pick 6 values with replacement, where 1 has a 60% chance of being picked and 0 has a 40% chance of being picked.  What about the sum of these 6 values tells you whether has a strike has occurred or not?

```{r Q6}


```