---
title: "BAS 471 Spring 2023 Homework on Unit 3 - Conditional Probability and Bayes' Theorem"
author: "Your Name Here Please"
date: "see Canvas for due date"
output: word_document
---

```{r setup, include=FALSE}
#Leave this as-is.
knitr::opts_chunk$set(echo = TRUE,collapse=TRUE)
if( as.numeric( paste( unlist( strsplit( paste(R.Version()$major,R.Version()$minor,sep="."),split="\\.") ),collapse="") ) < 422 ) { 
  stop("You must have R version 4.2.2 or later to knit this document")}
RNGversion("4.2.2")
```


******************************************************************************

# Note: these are your homework problems

As a reminder about the collaboration policy, remember that you can develop a common set of R code with your colleagues. However, all written interpretations (i.e., anything that follows a **Response:**) needs to be written *in your own words* (unless you are instructed to copy/paste what ChatGPT says for a particular question).  Homework submissions that look to be near copies of each other will receive substantially reduced credit, may receive a 0, and may result in you receiving an F in the class.  If you do use ChatGPT as a learning tool (caution:  in BAS 471, unlike other BAS classes, *there are quite a few concepts/problems that ChatGPT gets plain wrong*), you'll need to rephrase its very obvious AI-sounding phrasing.  Remember that exams are closed ChatGPT, so do not rely on it too much.  

******************************************************************************

For these questions, you'll be asked to compute conditional probabilities using the table method or using the laws of conditional probability (including Bayes Theorem).  Tables and equations are not the easiest to write in an .Rmd file.  If you wish to hand-write anything, or to do tables in Excel, that's fine!  You can take a screenshot or photo with your phone and paste them into the Word document that is produced *after* you knit.

**********************************************************************************

## Question 1 - Car Design

New cars can have a variety of cool features. Imagine a company that sells cars with the following characteristics:

- Some cars have a front dash cam, some have a rear dash cam, and some have neither.
- 25% of cars have both front *and* rear dash cams (and perhaps other cool features as well).
- 60% of cars have a front dash cam (and perhaps other cool features as well).
- 32% of cars have a rear dash cam (and perhaps other cool features as well).

a)  What is the probability that a car has a front or a rear dash cam? Think back to Unit 1.

**Response:**  
The probability of both of the cameras together is equal to the probability of a front camera, plus the probability of a rear camera, less the probability of both cameras. Since we already have all of this information in the prompt, we can write out the formula below.
P(Front U Rear) = P(Front)+P(Rear)- P(Front I Rear)= .6 + .32 -.25= .67

b)  What is the probability that a car has a front dash cam, given that it has a rear dash cam? In other words, what is P(front∣rear)? Reason out your answer without relying on the table method or Bayes' Theorem.

**Response:**  
Well we know that the formula for conditional probability is pretty useful here. Given that we know the intersect, which is 25%, we just need to know our sample space, which is the probability of the rear camera or 32%. Thus, the probability of having a front camera given a rear camera is our intersect over our sample space, or .25/.32 or 78.125%. 
I don't really know how to explain it without the formula, but the concept is that we change our sample space from the universal set to the set of interest or in this case the all the cars that have a rear camera. And once we know the proportion of all the cars that have a rear camera, we want to know what portion of them have a front camera as well, and that's our intersect. So yeah. 
Below is my formulaic approach to this. I represents intersect. 

P(front∣rear)= P(Front I Rear)/P(Rear) = .25/.32= 0.78125


```{r Q1 if you need it}
.25/.32



```

**********************************************************************************

## Question 2 - Sports Analytics Interview Questions

These questions are "actual" (slightly adapted) questions that BAS majors have been asked when interviewing for analytics positions for Major League Baseball Teams.

a)  From the Milwaukee Brewers:

Whether a runner on first base attempts to steal second base can depend on whether they are playing through an injury. Most players (92%) are fully healthy and have no injuries.

Imagine that a runner on first base will attempt to steal second base:

* 21% of the time when they are fully healthy.
* 6% of the time when they are playing through an injury.

This question will help you find the probability that, in general, a runner is playing through an injury when he attempts to steal second base.  

a1)  In general, what is the probability that the runner is playing through an injury?  Justify your answer.

**Response:** 

Around 8% of the time because 92% of players are healthy, and 1-.92= .08. I used the compliment rule here. 


a2)  Use the Table Method (with 5000 total runners) to answer the original question:  given that the runner attempts to steal second base, what is the probability that the runner is playing through an injury (sanity check:  a little larger than 2%)?  Fill in *all* cells of the table, not just the ones you need for the appropriate calculation.   Do your work by hand on paper (or in Excel), take a picture with your phone (or screenshot), then after knitting, insert that picture into your Word document.  Have your final computation in the Response.  You must use the Table Method to get credit for this question, though I recommend running the question through ChatGPT so you see the answer with Bayes Theorem as well.

Tips:

* Fill in the row totals first (very last column).  If you examine all 5000 runners, how many do you expect to be injured?  How many do you expect to be healthy?
* Fill in the column for "Steal".  Of the people who are healthy, how many are expected to steal?  Of the people who are injured, how many are expected to steal?  
* Fill in the column for "Not Steal".  Easy since the numbers in a row have to add up to the row totals you filled out first!
* Of all the times the runner attempts to steal, what fraction correspond to the runner being injured?  That's your conditional probability!


**Response:** 

 &nbsp;     | Not Steal   | Steal     | Total 
   -----    | ------      | ------    | ----
**Injured** |      376    |  24       |  400
**Healthy** |  3634       |     966   | 4600
**Total**   |   4010      |    990    | 5000

If you examine all 5000 runners, how many do you expect to be injured?  
400

How many do you expect to be healthy?
4600

Of the people who are healthy, how many are expected to steal?  
966

Of the people who are injured, how many are expected to steal? 
24

Of all the times the runner attempts to steal, what fraction correspond to the runner being injured? 
24/990=.02424

a3)  Summarize by filling in the blanks:

**Response:** In general, there is a 8% chance that a runner is injured.  Given that a runner attempts to steal second base, the probability he is injured decreases decreases by about a multiple of 3.3 (quote 1 digit after the decimal) to about 2.4% .   Note:  to get the "multiple", divide the prior and posterior probabilities appropriately so the ratio is greater than 1 (your multiple should be close to 3 here).


b)  From the Philadelphia Phillies:
 
Injuries factor a lot into gameplay.  A company has invented a device that turns red when exposed to an injured pitcher...most of the time (it stays green when exposed to a pitcher that is healthy).  The performance of the device is such that:

  * if the pitcher is injured (minor or severe), the device turns red 96% of the time
  * if the pitcher is healthy, the device turns green 90% of the time
  
At any given time, 14% of pitchers have some sort of injury (could be minor or could be severe).  Let's figure out the probability that a pitcher is injured if the device turns red.  

b1)  Use the Table Method to find P(injured|red).  Consider 10000 players.  Fill in all cells of the table, not just the ones you need for the appropriate calculation.   Have your final computation in the Response.  Sanity check:  a little below 61% (not a very useful device).  You must use the Table Method for this question to get credit, however I recommend running this question through ChatGPT to see how it is solved using Bayes Theorem.




**Response:**  Do your work by hand on paper (or in Excel), take a picture with your phone (or screenshot), then after knitting, insert that picture into your Word document.  You'll have to label the rows/columns yourself.


 &nbsp;     |  Red  |   Green |   Total 
   -----    | ----  | ----    | ----
    Injured | 1344  |   56    |   1400 
    Healthy | 860   | 7740    |   8600
   Total    | 2204  | 7796    |  10000
   
P(Injured|Red)= P(Injured I Red)/P(Red)= .1344/.2204= .6098

b2)  Fill in the blanks:  In general, there is a 14% chance that a pitcher is injured.  Given that the device turns red, this probability increases  by a multiple of about 4.36 (quote 3 digits total) to 60.9% .  Note:  to get the "multiple", divide the prior and posterior probabilities appropriately so the ratio is greater than 1 (your multiple should be close to 4.4 here).


c)  Bonus.  From the Tampa Bay Rays.  

Stan the stochastic southpaw always obeys the following rules to determine his next pitch.

* Pitch 1 will be a fastball 85% of the time.
* Future pitches are determined by the pitch that immediately preceded it.
  - If the previous pitch was a fastball, the following pitch will also be a fastball 75% of the time 
  - If the previous pitch was not a fastball, the following pitch will be a fastball 30% of the time
  
You start daydreaming about what you're going to do with your first big paycheck after you get hired in analytics.  You miss the identity of Stan's first 3 pitches, but you did see that Pitch 4 *is* a fastball.  In general (obviously, not for this specific instance, since the probability would either be 0% or 100%), when Pitch 4 *is* a fastball, what's the probability that Pitch 1 was also a fastball?  Sanity check: very close 87% (sorry ChatGPT, your answer of 93.3% when I tested you the first time is quite a bit off, and the 89.2% when I tested you the second time still isn't right).

Hint: this problem is best solved by using the fact that a conditional probability calculation is just a standard probability calculation that uses a "reduced" sample space to reflect the information that's provided.  The "original" sample space is the set of all 16 possible sequences of four pitches:  FFFF, FFFN, FFNF, FFNN, etc. (where F is fastball, N is not fastball).  Determine the "reduced" sample space (Pitch 4 is F), find the probabilities of each sequence in it, then find the "fraction" of the reduced sample space that contains our event of interest (Pitch 1 is F).  Note:  no further hints will be given by instructors or TAs.

```{r Q2c bonus}

SS <- expand.grid(first=c("F","N"), second = c("F","N"), third = c("F","N"), fourth=c("F","N"))



```


**********************************************************************************

## Question 3 - The Lost Hiker  

A hiker has gone missing after taking one of three trails: A, B, or C. Based on the trails' popularity, the prior probabilities of the hiker choosing these trails are 0.10 (A), 0.70 (B), and 0.20 (C).   

Each trail has different terrain, which affects the probability of a successful search:

* P(search A successful | hiker on A) = 0.90; A is not very popular (flat, monotonous, not many views), but searching it is easy
* P(search B successful | hiker on B) = 0.15; B is the most popular (gorgeous), but it's steep with cliffs, making searches hard
* P(search C successful | hiker on C) = 0.65; C 

The search and rescue team is preparing to begin their search. They suggest starting with Trail B because it’s the most popular. However, we’ll use data-driven reasoning to determine the best trail to search first.  This is one of the best questions of the semester because there is likely a disagreement between what your intuition suggests to search and what analytics says to search! 

a)  The search team says they'd like to start with trail B since it's the most likely to be chosen by the hiker (though it is the hardest to search).  Show that trail B is NOT the best choice for the first search.  Do so by finding the probability that a search on A would be successful, that a search on B would be successful, and that a search on C would be successful (13%).  To calculate each, you'll need to partition on the three possible locations of the hiker.  You should find that C is actually the best trail to start the search! 

```{r Q3a}



#P(search A successful) =.09
.9*.1


#P(search B successful) =.105
.15*.7

#P(search C successful) =.13
.65*.2
```

b)  The search team disregards your recommendation and decides to search Trail B first. Unfortunately, the search is unsuccessful—either because the hiker isn't on Trail B or the hiker is there but remains undetected due to the challenging terrain.  After this failed attempt, the team is now willing to follow your data-driven approach and let you decide the next trail to search.

Given that the search along B was unsuccessful, what are the posterior probabilities that the hiker is on trail A (close to 11%)?  B (66.48045%)?  C (close to 22%)?  

For this question, use the table method (Bayes Theorem approach gets *particularly* notation heavy).  Do your work by hand on paper (or in Excel), take a picture with your phone (or screenshot), then after knitting, insert that picture into your Word document.  Be sure to include computations that show your answers.

For the table method, use 8000 hikers and label the rows "Hiker on A", "Hiker on B", "Hiker on C" (our events of interest) and the columns "search B fails" and "search B succeeds" (the information on which we are conditioning the probability).  You only need to fill in the cells for "search B fails" (you may leave the other column blank for this and all further parts).  

ChatGPT will use Bayes Theorem here, though not always successfully.  The first time I asked it, it was able to math its way to the correct answer.  The second time I asked it (after asking it to forget all previous prompts), it made a mistake in logic that led it to a very incorrect answer.  

**Response:**  Place for you to copy/paste your work.  The R chunk below is for you to calculate the numerical values of the answers.

 &nbsp;              | search B fails  | search B succeeds | Total  
   -----             |-----------------|-------------------|-------
**Hiker on A**       |        800      |          0.       | 800
**Hiker on B**       |      4760       |        840.       | 5600
**Hiker on C**       |     1600        |          0        | 1600
**Total**            |      7160       |          840      | 8000

```{r Q3b}

HB<-.7*8000
HBS<-HB*.15
HC<-.2*8000
HA<-.1*8000
HBNS<-HB*(1-.15)

TBF<-HBNS+HC+HA

print(HA/TBF)
#Hiker on A given Trail B search fail
print(HC/TBF)
#Hiker on C given Trail B search fail
print(HBNS/TBF)
#Hiker on B, given Trail search B fail

```

c)  Your work in (b) shows that even if the initial search on B were to fail, the hiker is *still* most likely on trail B (in fact, about 3x as likely as the second most likely trail).  This is where "intuition" and analytics likely diverge.  If you asked the search team where they should search if an initial search on B were to fail, they'd likely say trail C (the 2nd most likely trail to be taken by hikers).  

However, if you stop and think about it, having trail B *still* be the most likely location does make sense.  B was by far the most likely trail taken by the hiker at 70%.  The fact that the search on B failed (even when the hiker is there, there's a 85% chance of failure) has diminished the probability that the hiker is on B down to about 66%, but this probability is *still* higher than the hiker being on C (the "distant" second most-likely trail for the hiker to have taken).

Thus, you instruct the search team to search on B again.  If both searches on trail B were to both fail, what are the posterior probabilities that the hiker is somewhere on trail A (about 12%)?  B (62.76761%)?  C (about 25%)?  What trail needs to be searched next? 

Hint:  you only need to make 3 changes to your table from (b).  The columns should now be labeled "search B fails twice" and "opposite" (i.e., one of the previous searches succeeded).  The only numbers in the "search B fails twice" column that will be different from your "search B fails" column in (b) is the cell for the row "Hiker on B" and the "Total".  *Do not* fill in the numbers of the column labeled "opposite".  ChatGPT (with a lot of coaxing, some re-wording, and some hand-holding) was able to get to this answer using Bayes Theorem, but doing this problem manually with the table method provides a much more sure-fire way to get to the answer as well as additional clarity). 

**Response:**  Place for you to copy/paste your work.  The R chunk below is for you to calculate the numerical values of the answers.  *Do comment on where we should search next*!

 &nbsp;              | search B fails twice  | opposite    | Total 
   -----             | ----------------------| ------------| ----
**Hiker on A**       |    800                |             | 800
**Hiker on B**       |       4046            |             | 4760
**Hiker on C**       |        1600           |    .        | 1600
**Total**            |          6446         |             | 7160



```{r Q3c}
#opposite column is blank because is what the prompt instructed, I prefer to not have it like that but yeah.

HB<-.7*8000
#OG number of times on trail B, not used in this calc much
HBS<-HB*.15
# number of times on trail B that search succeeds, not used much
HC<-.2*8000
#OG number of times on trail C, used in calculation
HA<-.1*8000
#number of times on trail A, used in calculation
HBNS<-HB*(1-.15)
#Fail search trail b, used here

TBF<-HBNS+HC+HA
#total fail B total from previous calculation, reduced sample space, is used as total/ total here


HBNS2<-HBNS*.85
#fail 2x on search of trail B
HBS2<-HBNS*.15
# success on 2x search of trail B

TSB2XF<-HBNS2+HC+HA
#total fail 2x, used for conditional probs
TTSB2X<-TSB2XF+HBS2
#total total, used for checking for accounting error, if i messed up a calculation. it returns 7160, I'm good. 
print(HC/TSB2XF)
#prob on trail C given search B failed 2x
print(HBNS2/TSB2XF)
#prob on trail B given search B failed 2x
print(HA/TSB2XF)
#prob on trail A given search B failed 2x
```

d)  Bonus (challenging).  Analytics really shines in this situation because it can tell us exactly what the search sequence should be in order to maximize the probability of finding the hiker.  Without analytics, you can imagine the strategy would have just been "let's search B, then C, then A".  Part (c) and establishes that the first three searches should be on trail B. 

* Imagine the first 3 searches on B all fail.  Where should the rescue team look next (on the 4th search).  If that fails, where should they look on the 5th search?  Etc.  

* Provide an optimal sequence of trails for the first 15 searches.  In this sequence, when does trail C finally get searched?  When does trail A finally get searched?   If you want to push yourself, try using a "for" loop in R to go well beyond the 15th search, reporting the first 50 searches!

*  Compare the probability of finding the hiker if the searchers follow your recommendation for the first 15 searches with the probability of searching BCA BCA BCA BCA BCA (what intuition probably suggests the search order should be).

**Response:** 

```{r Q3d}

```

**********************************************************************************

## Question 4 - Customer analytics

Retailers benefit greatly from identifying a new customer’s "type" as early as possible. Key questions include:

- Will this person be a one-time purchaser (a single purchase with no repeat visits)?
- Will this person be a casual purchaser (making repeat purchases every few weeks or months)?
- Will this person be a die-hard purchaser (making repeat purchases every few days)?

Once a retailer knows a customer's type, they can design personalized engagement and loyalty strategies to maximize retention and satisfaction.

Historically, at this particular particular retailer, 70% of customers are one-time purchasers, 20% are casuals, and 10% are die-hards.  The analytics team is building a classification model to predict a customer's type based on the monetary amount of their first purchase. Historical trends related to purchase behavior are summarized in the following table.  Historical trends are summarized in the following table (note, this is *different* from the table you'll use in the "table method" for calculating probabilities).

 &nbsp;  | 20 or less | 21-50 | 51-100 | 101 or more
 --------| ------     | ----  | ----   | ----------
One-time | 63%        | 17%   |  12%   |     8%
Casual   | 45%        | 27%   |  15%   |    13%
Die-hard | 26%        | 32%   |  22%   |    20%

The summary table shows that if a customer is a Casual purchaser, there's a 45% chance they spend 20 dollars or less on their first purchase, a 27% chance of spending between 21-50 dollars, a 15% chance of spending 51 or more dollars, and a 13% chance of spending 101 or more.

Let's focus on identifying die-hard customers.  Early identification of die-hard customers can have a significant impact on long-term revenue. If the retailer can identify these high-value customers early, it can implement strategies to incentivize future purchases with personalized offers, encourage more frequent visits through loyalty programs, and increase spending per visit by highlighting relevant, high-value products.  By tailoring engagement efforts, the retailer may increase both the frequency of visits and the average purchase amount beyond what the customer would have spent otherwise.

a)  In general (i.e., with no information about their first purchase), what is the probability a customer is a die-hard?

**Response:**  

10% of customers are die hard, as per the prompt. 

b)  Show that, in general (i.e., without knowing the customer's purchasing class),  the probability that a customer will spend 101 or more dollars on their first purchase is 10.2%.  Hint:  partition (or complete the table with the Table Method and calculate from there).

**Response:**  
10.2%, I used the partition method, it is in the chunk below. 


```{r Q4b}
omasot<-.08*.7
omasc<-.13*.2
omasdh<-.2*.1
pomas<-omasot+omasc+omasdh
print(pomas)



```

c)  Consider customers that spend 100+ on their first visit.  Die-hard customers are the most likely class to spend this amount.  However, die-hard customers are also the least common at the company.  You might wonder if the most likely identity of customers whose first purchase is 100+ is indeed die-hard customers.  

Determine the probability that a customer is a one-time customer (about 54.9%), a casual customer (about 25.5%), and a die-hard (about 19.6%) given that 100+ was spent on the first visit.  It turns out the most likely identity is *still* a one-time customer! 

To answer this question, use the method of your choice to find the posterior probabilities given they spend 100+ dollars on their first visit.  If you use the table method, consider 1000 customers.  You'll see a similar problem on the conceptual midterm, so make sure you're a master at working through the calculation!

**Response:**

  &nbsp;     | 20 or less | 21-50 | 51-100 | 101 or more  | Total
 --------    | ------     | ----  | ----   | ----------   | -----
One-time     |   441      | 119   |   84   |     56       |   700  
Casual       |    90      |   54  |    30  |    26        |    200
Die-hard     |    26      | 32.   |    22  |      20      |   100 
**Total**    |    557     |   205 |   136. |   102        |  1000 


```{r Q4c}
#101 or more

pomas*1000
omasot<-.08*.7
omasc<-.13*.2
omasdh<-.2*.1
omasc*1000
omasdh*1000
omasot*1000


print(omasot/pomas)
#probability that customer is one time given purchase of 101$ or more

print(omasc/pomas)
#probability that customer is casual given purchase of 101$ or more

print(omasdh/pomas)
#probability that customer is die hard given purchase of 101$ or more
0.1960784/.1
#total num in sample space
DH<- .1* 1000
# die hards
CAS<- .2 *1000
# casuals
OT<- .7* 1000
# one times


#51- 100
OT*.12
CAS*.15
DH*.22
(OT*.12)+(CAS*.15)+(DH*.22)
#21- 50
OT*.17
CAS*.27
DH*.32
(OT*.17)+(CAS*.27)+(DH*.32)

#20 or less
OT*.63
CAS*.45
DH*.26

(OT*.63)+(CAS*.45)+(DH*.26)



```

d) Fill in the blanks.  Initially, the probability a new customer is "die-hard" is 10%_.  However, knowing that a customer has spent 101+ dollars on their first purchase increases this probability by a multiple of 1.96  to 19.61%__.  Note:  to get the "multiple", divide the prior and posterior probabilities appropriately so the ratio is greater than 1 (your multiple should be very close to 2 here).

**Response:**  


e)  Comment on whether finding that a customer has spent 101+ dollars is sufficient for the analytics team to classify them as a die-hard (since die-hards are the most likely of the three classes to spend this amount).  If you don't think it is, and thus think more information is needed before such a classification can take place, propose something else that could be measured about a customer's first visit that could potentially prove useful.

**Response:** 

I don't think that is sufficient, because 54.9% of over $101 purchases, for first time pruchases, are from the group called one time shoppers. I suggest that the company has some sort of signaling system like a rewards program in order for the die hard customers to expose themselves as such, and I wouldn't push it super hard, like a simple question like "would you like to hear about our rewards programme" and wait for the customer's response. If it is a no, then they are could be a casual or a one time shopper, otherwise they could be a casual or die hard shopper. The problem would be that intersection with casual shoppers falling into both categories, however, if it's simply an email list this might be a better indicator of a customer's intentions than determining their type based off of their first purchase at a store. 



**********************************************************************************

## Question 5 (bonus)

The first day of class, one of the Kahoot questions was:

A driver made a dangerous maneuver on Kingston Pike.  The most likely identity of this driver is:

a)  Male who attends UT
b)  Female aged 20-45
c)  Student driver (learning with instructor)
d)  someone of Asian ethnicity

This question was designed to make you think about stereotypes and to wonder how much (if any, since stereotypes aren't always true) those factor into the answer.  While it's difficult to track down concrete numbers (try asking ChatGPT), the rates at which people in these four groups make dangerous maneuvers isn't *that* different from each other, perhaps varying by at most a factor of 2 (the highest rate is no more than double the lowest rate).  Do some research into Knoxville demographics, then determine the most likely identify of the driver, fully justifying your answer.  Hint:  write out Bayes Theorem -- this question is more easily answered looking at equations than tables.

**Response:**  
