---
title: "BAS 471 - Unit 9 - Part 6 - Bootstrapping"
author: "Boisterous Boostrappers"
output:
  word_document: default
---

```{r setup, include=FALSE}
#Leave this as-is.
knitr::opts_chunk$set(echo = TRUE,collapse=TRUE)
if( as.numeric( paste( unlist( strsplit( paste(R.Version()$major,R.Version()$minor,sep="."),split="\\.") ),collapse="") ) <= 400 ) { 
  stop("You must have R version 4.1.1 or later to knit this document")}
RNGversion("4.1.1")

install.packages("boot")
library(boot)
library(regclass)
```



*********************************************************************************************

# Note:  these are problems that are addressed in class and some additional practice problems

*********************************************************************************************


## Question 1 - bootstrapping with data coming from a Weibull 

Consider the following vector of data values `sales`, which is generated by a Weibull distribution with parameter.  These could be the sales (in units of thousands of dollars) of 45 B2B transactions.

```{r Q1} 
set.seed(471); sales <- rweibull(45,shape=.7,scale=3)
head(sales)
#26.7452283  8.5296701  0.5840896  0.2949704  1.2987073  3.3463835
hist(sales)
```

The Central Limit Theorem is not in play due to the extreme skewness of the distribution (with extreme skewness, we need at least 100 observations for the CLT to be in effect and we only have 45 here).  Thus, we do not know what the sampling distribution of possible sample averages that might arise when 45 data values are taken from the Weibull.

a)  Use a `for` loop to "manually" make a 95% percentile bootstrap confidence interval for "mu", the expected value of whatever probability distribution is making this data (which we happen to know in this case, but normally we don't).  Use 4999 bootstrap samples.  Compare with the result of running `t.test`.  Have `set.seed(471);` be on the same line but immediately before your `for` loop.   

Include a histogram of the averages of each bootstrap sample - you'll notice it doesn't quite look Normal, confirming that the CLT is not in effect (the CLT says the distribution of possible averages will be a Normal distribution if we consider samples that are "large enough").

**Kahoot 1**  How do you make a bootstrap sample from the values contained in `sales`?

```{r Q1a}
estimates <- c()
set.seed(471); for (i in 1:4999) {
  estimates[i] <- mean( sample(sales,replace=TRUE) )
}
hist(estimates)
quantile(estimates,c(.025,.975))
t.test(sales,conf=0.95)
```

b)  **Kahoot 2** Use the `boot` command to make the percentile and BCa confidence interval.  Notice how much different these two intervals actually are!

```{r Q1b}
library(boot)
estimator <- function(x,indices) { mean(x[indices]) }
set.seed(471); BOOT <- boot(sales,estimator,R=4999)
boot.ci(BOOT,type=c("perc","bca"),conf=0.95)
```

c)  Using the parameters of the Weibull that made this data (`shape=.7`,`scale=3`), find the actual expected value of the probability distribution making the data (it's close to 3.8).  Did your interval in (a) get it right and cover it?  Hint:  if look up ?rweibull to see the expected value, its formula is `scale*gamma(1+1/shape)`

**Response:*  Yes.

```{r Q1c}
shape <- .7
scale <- 3
scale*gamma(1+1/shape)
```

d)  **Kahoot 3**  While the bootstrap is relatively easy to program and doesn't require any math/statistical theory to compute confidence intervals, a shortcoming is that the achieved level of confidence is often lower than what's desired.

Consider 500 different random samples, each generated via `values <- rweibull(45,shape=.7,scale=3)` (with no random number seed).  Find 95% confidence intervals for "mu" using both the percentile and BCa methods (using only 499 bootstrap samples instead of 4999 due to time constraints), and estimate their achieved level of confidence.  Is it the desired level of 95%?  Use `binom.test` to get a 95% confidence interval on the achieved level of confidence.

Note 1:  if you store the output of `boot.ci` into an object called `intervals`, then `intervals$percent[4]` and `intervals$percent[5]` give the lower/upper bounds of the percentile bootstrap confidence interval (respectively), and analogously for `intervals$bca[4]` and `intervals$bca[5]`

Note 2:  you can ignore the warnings "In norm.inter(t, adj.alpha) : extreme order statistics used as endpoints", it's just letting you know the maximum value found in a bootstrap sample is used in the confidence interval calculations (and that's ok).

```{r Q1d achieved}
truth <- 3.797471

cover.perc <- 0
cover.bca <- 0
for ( trials in 1:5000) { 
  values <- rweibull(45,shape=.7,scale=3)
  BOOT <- boot(values,estimator,R=499)
  intervals <- boot.ci(BOOT,type=c("perc","bca"),conf=0.95)
  if( intervals$percent[4] <= truth & truth <= intervals$percent[5] ) {
    cover.perc <- cover.perc + 1
  }
  if( intervals$bca[4] <= truth & truth <= intervals$bca[5] ) {
    cover.bca <- cover.bca + 1
  }
}
cover.perc/5000
cover.bca/5000

#Noticeably lower than the target 95% level of confidence
binom.test(cover.perc,5000)
binom.test(cover.bca,5000)
```

e)  **Kahoot 4** Let's say we did not know this data was coming from a Weibull.  Could we use the bootstrap to obtain a 95% confidence interval of the probability that a sale would obtain a value between 6.1 and 7.7?   Hint:  look at `sort(sales)`

**Response:**  No.  Since no such values exists in the original data, no such values will exist in the bootstrap samples.  Each bootstrap sample would have an estimated probability of 0.  But surely there is some probability of this occurring.

f)  **Kahoot 5** It is desired to come up with an approximate 95% confidence interval for the quantity:  "80th percentile" - 2*"20th percentile".  On the original data values, this works out to be:

```{r Q1f illustrate}
quantile(sales,.8)-2*quantile(sales,.2) 
```

Use the BCA method with 4999 bootstrap samples.  Is 9.2 a plausible value for this quantity?

```{r Q1f problem}
estimator <- function(x,indices) { quantile(x[indices],.8)-2*quantile(x[indices],.2)  }
set.seed(471); BOOT <- boot(sales,estimator,R=4999)
boot.ci(BOOT,type="bca")
```



## Question 2 - bootstrapping using the Wine data.

The bootstrap is a powerful tool that allows us produce a confidence interval for *almost any* parameter of the probability distribution responsible for producing our data, often without even knowing exactly what that distribution is.   Consider the `WINE` dataset, which records chemical properties of high quality and low quality wine.  If we had a full understanding of what characteristics (individual and combinations of) were associated with being high quality, we could use analytics to help design a "new" high quality wine!

a.  Left-arrow into a vector called `chlorides.lowQ` the values in the `chlorides` column for `low` quality wines (a `summary` of this vector is provided as a sanity check).  A histogram of data values reveals it is quite skewed.  When this is the case, the median serves as a better summary of the "typical" value than the mean.  

a1.  Using the `boot` package, produce a 95% confidence interval (percentile method) for the median value of `chlorides` for `low` quality wines.  Use 4999 bootstrap samples and have `set.seed(471)` on the same line and immediately preceding the `boot` command.

a2.  Also produce a 95% confidence interval for the 80th percentile. Use 4999 bootstrap samples and have `set.seed(471)` on the same line and immediately preceding the `boot` command.

```{r Q2a}
library(regclass)
data(WINE)
chlorides.lowQ <- WINE$chlorides[which(WINE$Quality=="low")]

# summary(chlorides.lowQ)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#0.00900 0.04000 0.04700 0.05144 0.05300 0.34600 


library(boot)
estimator <- function(x,indices) { return( median(x[indices]) ) }
set.seed(471); BOOT <- boot(chlorides.lowQ,estimator,R=4999)
boot.ci(BOOT,type="perc",conf=0.95)
```


b.  Also produce a 95% confidence interval for the 80th percentile. Use 4999 bootstrap samples and have `set.seed(471)` on the same line and immediately preceding the `boot` command.

```{r Q2b}
estimator <- function(x,indices) { return( quantile(x[indices],0.80) ) }
set.seed(471); BOOT <- boot(chlorides.lowQ,estimator,R=4999)
boot.ci(BOOT,type="perc",conf=0.95)
```



c.  Left-arrow into a vector called `sugar.highQ` the values in the `residual.sugar` column for `high` quality wines (a `summary` of this vector is provided as a sanity check).  Produce the standard "classic" 95% confidence interval for the mean using `t.test` and also with the bootstrap (both percentile and BCa methods).  Assuming each procedure achieves 95% confidence, which of these procedures is "better" (among intervals with the same level of confidence, skinnier is better than wider).  Use 4999 bootstrap samples and have `set.seed(471)` on the same line and immediately preceding the `boot` command.

```{r Q2c}
sugar.highQ <- WINE$residual.sugar[which(WINE$Quality=="high")]
summary(sugar.highQ)
  #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  # 0.800   1.800   3.875   5.262   7.400  19.250 
t.test(sugar.highQ)
estimator <- function(x,indices) { return( mean(x[indices]) ) }
BOOT <- boot(sugar.highQ,estimator,R=4999)
boot.ci(BOOT,type=c("perc","bca"),conf=0.95)

```

**Response:**  A confidence interval for the average is `t.test`s playground, but the BCa method in this case provides a slightly skinnier (and thus better) interval with a width of 0.517 (t.test) vs. 0.51 (BCa).  The bootstrap is a much more flexible procedure.



d.  **Kahoot 6**  Of interest is the alcohol content where a wine has a 50-50 chance of being high vs. low.  Transporting us back to BAS 320, this can be found by fitting a logistic regression to the quality with alcohol as a predictor.  The relationship between these two look roughly described by the logistic curve that the model imposes, so let's go with it.  The estimated alcohol content is the ratio of the "intercept" to the coefficient of alcohol (then flipping the sign), which we can see is just about 10.9 here.  

Let's write the `estimator` function so that the argument `x` is the dataframe of the Wine data.  This way we can refer to the bootstrap dataset as `x[indices,]` (`indices` refers to the rows selected for the bootstrap sample).

Obtain a 95% confidence interval for the alcohol content associated with a 50-50 chance of being high vs. low quality using 500 bootstrap samples.   Use the percentile method (again, BCa bugs out for some weird reason).

```{r Q2d}
plot(Quality ~ alcohol,data=WINE)
M <- glm(Quality ~ alcohol, data=WINE, family="binomial")
visualize_model(M)
fifty.fifty <- -coef(M)[1]/coef(M)[2]
abline(v=fifty.fifty)

even.chance <- c()
n <- nrow(WINE)
for (i in 1:999) {
  M <- glm(Quality ~ alcohol, family="binomial", data=WINE[sample(1:n,size=n,replace=TRUE), ])
  even.chance[i] <- -coef(M)[1]/coef(M)[2]
}
quantile(even.chance,c(.025,.975) )

estimator <- function(x,indices) {
  M <- glm(Quality ~ alcohol, data=x[indices,], family="binomial")
  fifty.fifty <- -coef(M)[1]/coef(M)[2]
  return(fifty.fifty)
}
BOOT <- boot(WINE,estimator,R=499)
boot.ci(BOOT,type=c("perc"),conf=0.95)

```








