---
title: "BAS 471 - Unit 8 - Part 3 and Part 9 intro - Bias, Variance, and RMSE of Estimators"
author: "You are the best estimator ever!"
date: "see Canvas for due date"
output:
  word_document: default
---


```{r setup, include=FALSE}
#Leave this as-is.
knitr::opts_chunk$set(echo = TRUE,collapse=TRUE)
if( as.numeric( paste( unlist( strsplit( paste(R.Version()$major,R.Version()$minor,sep="."),split="\\.") ),collapse="") ) <= 400 ) { 
  stop("You must have R version 4.1.1 or later to knit this document")}
RNGversion("4.1.1")
```

*********************************************************************************************

# Note:  these are problems that are addressed in class and some additional practice problems

*********************************************************************************************


## Question 1:  Maximum Likelihoood for Censored Data

The time to failure of a machine part is exponential with parameter $\lambda$

$$f(x) = \lambda e^{-\lambda x} \quad x>0$$
$$F(x) = P(X \le x) = 1 - e^{-\lambda x}$$
On a previous worksheet, the maximum likelihood estimator of $\lambda$, given a random sample of data values and their average $\bar{X}$ is:

$$\hat{\lambda}_{MLE} = \frac{1}{\bar{X}}$$
However, imagine that when measuring the lifetimes of these parts, some lifetimes exceeded the duration of the experiment.  Specifically, consider an experiment that attempted to measure the lifetimes of 13 parts during an accelerated testing of 6 days, but the lifetimes of 5 of those parts were still functioning when the experiment ended (their exact lifetimes are unknown).

The following 8 lifetimes (in days) were observed

```{r Q1 intro}
lifetimes <- c(1.01, 1.62, 1.80, 1.84, 2.35, 2.89, 2.89, 3.66)
```

All that's known about the lifetimes of the other 5 parts is that they are greater than 6 days.

a)  Plugging $x=6$ into the CDF gives us $P(X \le 6)$.   Thus, in term of $\lambda$, what is the probability that the lifetime of a part is *greater* than 6?

**Response:** 


b)  **Kahoot 1** Write out the likelihood of the data (the probability of observing the 8 known lifetimes multiplied by the probability that 5 lifetimes are greater than 6).  Using Wolfram-Alpha, find the MLE estimate for $\lambda$.  

**Response:**  


***********

## Question 2 - bias, variance, root mean squared error

Previously, we talked about the Pareto distribution, a PDF often used to describe the distribution of wealth (and is the origin of the 80-20 rule, e.g., 80% of profit comes from 20% of customers). 

$$f(x) = \frac{3m^3}{x^4} \qquad x \ge m$$

In other words, the probability of finding larger and larger values of $x$ very quickly gets smaller.  All the while, there is an unknown minimum value that $x$ can attain (the parameter $m$).  You can show it is a valid PDF via `integrate 3*m^3/x^4 from x=m to x=infinity` in Wolfram-Alpha.

In our previous analysis, we argued that the MLE of $m$ was equal to the smallest value observed in our random sample (the likelihood gets larger as $m$ gets better, but $m$ can't be larger than any of the sample values).

$$\hat{m}_{MLE} = \min \{X_i\}$$

a)  Find the methods of moments estimator for $m$.  Specifically, provide a screenshot showing the expected value of the Pareto distribution, then another screenshot showing how you solved for the MoM.

**Response:** 

b)  The MLE and the MOM both provide reasonable ways to estimate $m$. To determine which is "best", we can look at their "root mean squared errors" (a combination of the estimator's bias and its variance), i.e. the "typical" error made by the estimator.  

For "large" sample, the MLE can't be beat, but for smaller samples it may not be the optimal estimator. What if we knew that a Pareto was generating our data, but we only had a sample of size 5?  Let's investigate.  So that we can quantify how "wrong" these estimators are, let's imagine that the "true" value of $m$ is $m=7$.

Run a Monte Carlo simulation (100,000 trials) to evaluate the bias, variance, and root mean squared error of both of these estimators for samples of size 5 when the "true" value of $m$ is 7.

* Define vectors `mle` and `mom` to be empty vectors before your `for` loop.

* In each trial, generate a random sample of size 5 (an example is provided).  Note:  there are clever ways to generate random samples from any distribution you have in mind.  That kind of material is covered in the graduate program.

* Measure the values of the MLE estimator and the MoM estimator on the sample, storing it into the appropriate position of your `mle` and `other` vectors.

* **Kahoot 2**  Find the bias of each estimator

* Find the variance of each estimator

* **Kahoot 3**  Find the root mean squared error (RMSE) of each estimator.

```{r Q2b}
#The following line of code will generate a random sample from the Pareto with m=7
x <- 7/(1-runif(5))^(1/3)

#Define mle and mom to be empty vectors
mle <- c()
mom <- c()

#Run Monte Carlo simulation.  Each trial generate a random sample then store into
#the mle and mom vectors the values of the MLE and MOM estimators

mle<-min(x)
mom<-2*mean(x)/3

mle <- c()
mom <- c()
for (trial in 1:10000){
  x <- 7/(1-runif(5))^(1/3)
  mle[trial]<-min(x)
  mom[trial]<-2*mean(x)/3
  
}

mean(mom)-7
mean(mle)-7

hist(mom,breaks=100,xlim=c(4,10))
hist(mle,breaks=100,xlim=c(4,10))

#Report the bias of each estimator (average value of the estimator minus the "true" value)
sd(mom)
sd(mle)

var(mom)
var(mle)

#Report the variance of each estimator  (the `var()` of values of the estimators)

sqrt((mean(mom)-7)^2 +var(mom))

sqrt((mean(mle)-7)^2 +var(mle))

#Report the RMSE of each estimator  (square root of bias^2 + variance)


```

c)  **Kahoot 4**  Fill in the blanks:  the typical error when the MLE is used to estimate $m$ is _____ while the typical error when the MOM is used to estimate $m$ is _____.  Which estimator is "better"?

**Response:**  


d)  **Kahoot 5**   When we are estimating the parameters of probability distributions and we have a lot of data, the maximum likelihood procedure is all you need because...

**Response::**  

e)  We've approached this problem mathematically/numerically so far.  Let's assess the quality of these two estimators graphically by looking at the sampling distributions of these two estimators and comparing their biases and variances.

The following code make a dataframe where the first column is the value of the estimator over the 100,000 samples created in the Monte Carlo simulation and the second column is the name of the estimator (this format allows for easy plotting).

* The horizontal red line represents the "true" value of $m=7$
* The horizontal line in the boxplot is the *median* value of the estimator over the 100,000 samples (ignore it, it doesn't factor into our analysis here)
* The blue dot is the *average* value of the estimator over the 100,000 samples and what we'll focus in on.

```{r Q2e,eval=FALSE}
#Will only run after the rest of the parts are completed
ESTIMATORS <- data.frame(Value=c(mle,mom),Estimator=factor(c(rep("MLE",length(mle)),rep("MOM",length(mom)))))
plot(Value ~ Estimator, data=ESTIMATORS,ylim=c(4,20))
abline(h=7,col="red",lwd=3)
points(1,mean(mle),pch=20,cex=3,col="blue")
points(2,mean(mom),pch=20,cex=3,col="blue")
```

Observations:

* The MLE has a *positive* bias.  The average value of the MLE estimator on a random sample (blue dot at 7.5) is well-above the red horizontal line giving the true value of $m$ (7).
* The MOM is *unbiased*.  The average value of the MOM estimator on a random sample (blue dot) is right at the red horizontal line giving the true value of $m$ (7).
* The MLE has a relatively *small* variance - the values of that estimator on random samples is relatively tightly clustered around its average (upper/lower whiskers of the boxplot).
* The MOM has a relatively *large* variance - the values of that estimator on random samples are spread loosely around its average (upper/lower whiskers of the boxplot).

Thus while the MLE is biased, it turns out to be a better estimator since it's "typical error" (RMSE = sqrt(bias^2+variance)) is smaller.



