---
title: "BAS 471 - Unit 5 Parts 4 and 5 - Expected Value, Standard Deviation, Checking Fit"
author: "Pontiffs of the PMF"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Question 1

Let the random variable X be the demand for a product.  Possible values are x=8 to x=46.

```{r Q1}
x <- 8:46
shape <- 6300 + 18*x - 0.012*(x - 19)^2*(x-74)^2
p<-shape/sum(shape)
mu<-sum(p*x)
var<-sum(((x-mu)^2)*p)
sqrt(var)
mu

barplot(shape,names.arg=x,xlab="Possible Demand",ylab="Relative Frequency")
```

a.  The values in `shape` are not probabilities, but we can convert them into them.  Use the standard shape to PMF conversion to define a vector `p` whose elements give the probabilities of observing each value of `x`.  Verify your output of running `head` on `p`, then calculate the probability that the demand is between 16 and 26 inclusive.


```{r Q1a}
sum(p[which(x<=26 & x>=16)])



#head(p)
#0.0008542915 0.0099856723 0.0179246261 0.0247475949 0.0305289548 0.0353410159


```

b.  Find the expected value of the demand.

```{r Q1b}

```

c.  What's the difference between the expected value, the average, and the mean of a probability distribution?

**Response:**  

d. What's the standard "layman's" interpretation of the numerical value of the expected value of a probability distribution?  In other words, how should we talk about it to the boss?

**Response:** 

e. Calculate the standard deviation of the demand.

```{r Q1e}

```

f. How do we interpret the standard deviation of a probability distribution?

**Response:**

g.  Provide an interpretation of the average and standard deviation of the distribution of demand in the usual boss-friendly way, i.e., in the context of the problem.

**Response:**  

h.  Imagine that the cost to ship a particular demand is:

$$Cost = 10 + 0.5\sqrt{x+2}$$
`cost = 10 + 0.5*sqrt(x+2)`

In other words, shipping cost increases, but cost per item diminishes for larger numbers of items.

Find the expected cost of a random day's demand (it's between 12 and 13 )

```{r Q1h}
Cost = 10 + 0.5*(sqrt(x+2))
Cost = 10 + 0.5*(sqrt(mu+2))

sum(Cost*p)

```

i.  One theme of probabilistic modeling is "averages aren't good enough".  If we want to characterize the typical value of a complex random process (like the cost of shipping some demand), it's not acceptable to plug in the average value of $X$ into our equation for Cost  What is the value you get if you plug in the average value of $X$ (from part b) into the equation for $Cost$?   In this case, it's actually pretty close to the actual average cost!  But it isn't always, as you'll see in the homework problems.


```{r Q1i}

```

*****************

## Question 2

Make sure this .Rmd file and the `FitTheseDistributions.RData` have been moved/saved to your BAS 471 folder.  

```{r Q2}
library(fitdistrplus)
load("FitTheseDistributions.RData")
```

For each vector:

* fit the data to the requested distribution (left-arrow it into an object with the name `FIT`)
* print the contents of `FIT` out to the screen to see the best-fitting parameters of the distribution
* include `plot(FIT)` to assess the fit
* comment on whether (and why) it's a good fit, bad fit, or otherwise 

**NOTE**:  when you run `fitdist`, some text will automatically be output to the screen.  These are back-end details about the procedure used to determine the best-fitting model and they are not relevant to us.  Ignore it!  The actual details about the fitted model is found by printing the contents of `FIT` to the screen.

a)  `x1.totalitems` - the total number of items in people's shopping carts at checkout (real data from Kroger actually).  Fit to a Poisson and to a Negative Binomial.  Which is better?  Do either provide a reasonable reflection of reality?

```{r Q2a}
FIT<-fitdist(x1.totalitems, dist="nbinom")
plot(FIT)
summary(FIT)

```

b)  `x2.totalvisits` - the total number of visits that households have made to Kroger over the course of 2 years.  Fit to a Poisson and to a Negative Binomial.  Which is better?  Do either provide a reasonable reflection of reality?

```{r Q2b}

```

c)  `x3.spampurchases` - the total number of times that households purchased Spam at Kroger over the course of 2 years.  Fit to a Poisson and to a Negative Binomial.  Which is better?  Do either provide a reasonable reflection of reality?

```{r Q2c}

```

d)  `x4.numberof16redeemed` - the number out of 16 coupons that a Kroger customer was sent that they ended up redeeming.  Code is provided to fit to a binomial (you have to specify an argument to that gives the n=size parameter and a starting "guess" of what the probability of success might be; the best-fitting Binomial may have a different number).

```{r Q4d}

```

e)  `x5.custservicecalls` - the number out times the customer called customer service.   Fit to a geometric, Poisson, binomial (with the provided code), and negative binomial.  Which provide reasonable reflections of reality?

```{r Q4e}

FIT <- fitdist(x5.custservicecalls, dist="nbinom") 


FIT <- fitdist(x5.custservicecalls, dist="geom")

FIT <- fitdist(x5.custservicecalls, dist="pois")
FIT <- fitdist(x5.custservicecalls, dist="binom", fix.arg=list(size=150), start=list(prob=2/150))
plot(FIT)
```

f)  `x6a.timebetweenvisits` - the number of days (can be 0) that elapsed between shopper A's visits to Kroger.  Fit to a geometric, Poisson, binomial (with the provided code), and negative binomial.  Which provide reasonable reflections of reality?

```{r Q4f}



FIT <- fitdist(x6a.timebetweenvisits, dist="binom", fix.arg=list(size=150), start=list(prob=12.8/150))

```


g)  `x6b.timebetweenvisits` - the number of days (can be 0) that elapsed between shopper B's visits to Kroger.  Fit to a geometric, Poisson, binomial (with the provided code), and negative binomial.  Which provide reasonable reflections of reality?

```{r Q4g}



FIT <- fitdist(x6b.timebetweenvisits, dist="binom", fix.arg=list(size=100), start=list(prob=.03))

```

h)  `x6c.timebetweenvisits` - the number of days (can be 0) that elapsed between shopper C's visits to Kroger.  Fit to a geometric, Poisson, and negative binomial.  Which provide reasonable reflections of reality?

```{r Q4h}
FIT <- fitdist(x6c.timebetweenvisits, "geom")
plot(FIT)  #decent
FIT <- fitdist(x6c.timebetweenvisits, "nbinom")
plot(FIT)  #decent
FIT <- fitdist(x6c.timebetweenvisits, "pois")
plot(FIT)   #bad
```

i)  `x6d.timebetweenvisits` - the number of days (can be 0) that elapsed between shopper D's visits to Kroger.  Fit to a geometric, Poisson, and negative binomial.  Which provide reasonable reflections of reality?

```{r Q4i}
FIT <- fitdist(x6d.timebetweenvisits, "geom")
plot(FIT)  #good
FIT <- fitdist(x6d.timebetweenvisits, "nbinom")
plot(FIT)  #good
FIT <- fitdist(x6d.timebetweenvisits, "pois")
plot(FIT)  #bad
```

j)  `x6f.timebetweenvisits` - the number of days (can be 0) that elapsed between shopper D's visits to Kroger.  Fit to a geometric, Poisson, and negative binomial.  Which provide reasonable reflections of reality?

```{r Q4j}
FIT <- fitdist(x6f.timebetweenvisits, "geom")
plot(FIT)  #good
FIT <- fitdist(x6f.timebetweenvisits, "nbinom")
plot(FIT)  #good
FIT <- fitdist(x6f.timebetweenvisits, "pois")
plot(FIT)  #not good
```

k)  `x8.harrypotter` - the number of Harry Potter books read by students.  Fit to a binomial distribution with n=7.  Is it a good fit?

```{r Q4k}
FIT <- fitdist(x8.harrypotter, dist="binom", fix.arg=list(size=7), start=list(prob=.46)); FIT
plot(FIT)  #terrible
```

l)  `x11.numcategories` - the number of different product categories (out of 6; e.g. electronics, clothing, grocery, etc.) in which consumers have made purchases.

```{r Q4l}
FIT <- fitdist(x11.numcategories, dist="binom", fix.arg=list(size=6), start=list(prob=.46)); FIT
plot(FIT)  #ok at best
```


m)  `x12.numdonations` - the number of donation a previous donor has given to a veteran's charity.  Fit to a geometric, Poisson, and negative binomial.  Which provide reasonable reflections of reality?

```{r Q4m}
FIT <- fitdist(x12.numdonations, dist="pois")
plot(FIT)  #ok but not great
FIT <- fitdist(x12.numdonations, dist="nbinom")
plot(FIT)  #decent
FIT <- fitdist(x12.numdonations, dist="geom")
plot(FIT)  #nope
```

