---
title: "Econometrics PS1"
author: "Grayson Pettit"
date: "2025-08-27"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Sets the working directory.
(b) Creates a log (perhaps using sink, or R Markdown).
(c) Inputs the following matrices:

(d) Finds AB′
(e) Tries to compute a non-conforming matrix product
(f) Uses an assertion statement to confirm that (AB′)′
= BA′
.
(g) Uses an assertion statement to confirm that tr(AB′C) = tr(D), where D is a cyclic permuta-
tion of AB′C.
(h) Finds A−1 using solve and uses an assertion statement to confirm that AA−1 = I.


```{r P6}

getwd()


X<-c(2,0,1,4,3,2,0,9,1)
A<-matrix(data = X, nrow = 3, ncol = 3, byrow = TRUE,dimnames = NULL)
Y<-c(0,0,1,1,1,0)
B<-matrix(data = Y, nrow = 2, ncol = 3, byrow = TRUE,dimnames = NULL)
Z<-c(4,9,0,7,4,8)
C<-matrix(data = Z, nrow = 2, ncol = 3, byrow = TRUE,dimnames = NULL)
print(A)
print(B)
print(C)

??transpose
AT<-t(A)
print(AT)
BT<-t(B)
print(BT)

ABT<-A %*% BT
BAT<- B %*% AT
print(BAT)
ABTT<-t(ABT)

print(ABTT)

ZERO<-BAT-ABTT
print(ZERO)


ABTC<-ABT %*% C
print(ABTC)
t(ABTC)

# Compute AB'C
expr1 <- A %*% t(B) %*% C

# Cyclic permutation: B'CA
expr2 <- t(B) %*% C %*% A

# Another cyclic permutation: CAB'
expr3 <- C %*% A %*% t(B)

# Compute traces
tr1 <- sum(diag(expr1))
tr2 <- sum(diag(expr2))
tr3 <- sum(diag(expr3))

print(tr1)
print(tr2)
print(tr3)

#tr1=tr2=tr3=136

A_inv= solve(A)
print(A_inv)

AIA<-A %*% A_inv

I<-diag(3)
print(I)
AIA-I







```

(i) Loads loanmain.csv and finds the mean and standard deviation of lnpart5revenue and
total_profit variables.
• Bonus: Finds the variable mean and standard deviation (1) formulaically and (2) using
commands, and uses an assertion statement to confirm that they are the same.


```{r P6loan, echo=FALSE}
library(readr)
library(fitdistrplus)
loanmain <- read_csv("loanmain.csv")

```

ln part 5 revenue 
Standard deviation and MU

```{r P6rev, echo=FALSE}

DATA<- as.numeric(na.omit(loanmain$lnpart5revenue))
SD<-sd(DATA)
SD # Standard Deviation
MU<-mean(DATA)
MU #Mean

TOTAL<-sum(DATA)
n<-length(DATA)

Mean=TOTAL/n
Mean
Mean-MU #This is basically 0.

StandardDEV<-sqrt(mean((DATA - mean(DATA))^2))
StandardDEV
SD-StandardDEV # this is also basically 0
```

Total Profit 
Standard deviation and MU

```{r}
DATA<- as.numeric(na.omit(loanmain$total_profit))
SD<-sd(DATA)
SD # Standard Deviation
MU<-mean(DATA)
MU #Mean

TOTAL<-sum(DATA)
n<-length(DATA)

Mean=TOTAL/n
Mean
Mean-MU #This is basically 0.

StandardDEV<-sqrt(mean((DATA - mean(DATA))^2))
StandardDEV
SD-StandardDEV # this is also basically 0
```

(j) Finds the mean of newloan by treatment status (given by type).
Mean of loans made that are new given type 0 is 6.7%
Mean of loans made that are new given type 1 is 34.7%
```{r}
table(loanmain$newloan)

aggregate(newloan~ type, data = loanmain, FUN= mean, na.rm=TRUE)

```

